---
title: "Games"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Packages

```{r Load Packages, message=FALSE}
library(dplyr)
library(ggplot2)
library(lubridate)
```

# Read In Data

```{r Read In Data}
data = read.csv('/Users/chengyinliu/D/Projects/Pet_Project/Games/Data/ign.csv', na.strings = '')
```

# EDA

```{r}
summary(data)
```

Year: from 1970 to 2016

We can see that there are Null Value in genre, so we delete them at the beginning

```{r}
data_c <- data %>%
  filter(!is.na(data$genre))
```

## Questions want to answer:

1. Which platforms are popular?

2. Which kinds of genre are popular?

3. When is the best time to release a game?

4. What are the main topic of games?

## Which platforms are popular?

```{r}
data_c %>%
  group_by(platform) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) %>%
  ggplot(mapping = aes(x = reorder(platform, count), 
                       y = count)) +
  geom_bar(stat = 'identity') +
  coord_flip()
```

### Top 20 

```{r}
data_c %>%
  group_by(platform) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) %>%
  head(20) %>%
  ggplot(mapping = aes(x = reorder(platform, count), 
                       y = count)) +
  geom_bar(stat = 'identity') +
  coord_flip()
```

### Top 20 percentage

```{r}
total_num = 18589
data_c %>%
  group_by(platform) %>%
  summarise(count = n()) %>%
  mutate(count_per = count / total_num) %>%
  arrange(desc(count_per)) %>%
  mutate(pre = '%') %>%
  mutate(label = as.character(round(count_per * 100, digits = 2))) %>%
  mutate(label_1 = paste0(label, pre)) %>%
  head(20) %>%
  ggplot(mapping = aes(x = reorder(platform, count_per), 
                       y = count_per)) +
  geom_bar(stat = 'identity') +
  geom_text(aes(label = label_1), 
            color = 'red', 
            vjust = .5, 
            hjust = -.3) +
  scale_y_continuous(limits = c(0,0.2)) +
  coord_flip()
rm(total_num)
```

Observation:

1. PC has the most games on it compared to other platforms, more than about twice as many as the second, which is PlayStation 2
2. 50% of the games supports the platforms of PC, PlayStation 2, Xbox 360, Wii, PlayStation 3

## Which kinds of genre are popular?

### For game maker

### Regarding Count

```{r}
data_c %>%
  group_by(genre) %>%
  summarise(count = n()) %>%
  ggplot(mapping = aes(x = reorder(genre,count),
                       y = count)) +
  geom_bar(stat = 'identity') +
  coord_flip() +
  theme(axis.text.y = element_text(size = 3))
```

### Top 20

```{r}
data_c %>%
  group_by(genre) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) %>%
  head(20) %>%
  ggplot(mapping = aes(x = reorder(genre,count),
                       y = count)) +
  geom_bar(stat = 'identity') +
  coord_flip() 
```

### Top 20 Percentage

```{r}
total_num = 18589
data_c %>%
  group_by(genre) %>%
  summarise(count = n()) %>%
  mutate(count_per = count / total_num) %>%
  arrange(desc(count_per)) %>%
  mutate(pre = '%') %>%
  mutate(label = as.character(round(count_per * 100, digits = 2))) %>%
  mutate(label_1 = paste0(label, pre)) %>%
  head(20) %>%
  ggplot(mapping = aes(x = reorder(genre,count_per),
                       y = count_per)) +
  geom_bar(stat = 'identity') +
  geom_text(aes(label = label_1), 
            color = 'red', 
            vjust = .5, 
            hjust = -.3) +
  scale_y_continuous(limits = c(0,0.25)) +
  coord_flip() 

rm(total_num)
```

Observation:

1. Action game is the most popular game for the game makers, 20% of the games are action game, almost more than twice as many as the second one.
2. 50% of the games made during 1970 to 2016 are Action, Sports, Shooter, Racing and Adventure games.

### For players

## Score Distribution

```{r Score}
data %>%
  ggplot(mapping = aes(x = score)) +
  geom_histogram(bins = 10,
                 color = 'black')
```

```{r}
data_c %>%
  group_by(genre) %>%
  summarise(avg_score = mean(score)) %>%
  ggplot(mapping = aes(x = reorder(genre,avg_score),
                       y = avg_score)) +
  geom_bar(stat = 'identity') +
  coord_flip() +
  theme(axis.text.y = element_text(size = 3))
```

### Top 20

```{r}
data_c %>%
  group_by(genre) %>%
  summarise(avg_score = mean(score)) %>%
  arrange(desc(avg_score)) %>%
  head(20) %>%
  ggplot(mapping = aes(x = reorder(genre,avg_score),
                       y = avg_score)) +
  geom_bar(stat = 'identity') +
  coord_flip() +
  geom_text(aes(label = sprintf('%.2f', avg_score)),
            color = 'red',
            vjust = .5, 
            hjust = -.3) +
  scale_y_continuous(limits = c(0,11))
```

Obervation:
1. Compilation has the highest score among all the game genres (this may because the game makers only choose the popular games together to make the compilation)


What about not popular games among players?

```{r}
data_c %>%
  group_by(genre) %>%
  summarise(avg_score = mean(score)) %>%
  arrange(desc(avg_score)) %>%
  tail(20) %>%
  ggplot(mapping = aes(x = reorder(genre,-avg_score),
                       y = avg_score)) +
  geom_bar(stat = 'identity') +
  coord_flip() +
  geom_text(aes(label = sprintf('%.2f', avg_score)),
            color = 'red',
            vjust = .5, 
            hjust = -.3) +
  scale_y_continuous(limits = c(0,11))
```

Observation:
1. It is interesting that Hunting Action games have the lowest score among the players, because as far as I know, games like Monster Hunter should be popular and the feedbacks from the players are almost positive, so maybe there are some other reasons behind it. 

### What Hunting Games do we have?
```{r}
data_c %>%
  filter(genre == 'Hunting, Action')
```

So this dataset doesn't consider Monster Hunter as a Hunting Game, so what the bottom 5 games?

```{r}
data_c %>%
  filter(genre %in% c('Hunting, Action', 'Sports, Fighting', 'Sports, Editor', 'Educational, Trivia', 'Adventure, Adult'))
```

# When is the best time to release a game?

```{r}
data_c$Date <- ymd(paste0(data_c$release_year, '-', sprintf('%02d', data_c$release_month), '-', sprintf('%02d', data_c$release_day)))
```

## Count

### Year

```{r}
data_c %>%
  mutate(year = year(Date)) %>%
  group_by(year) %>%
  summarise(count = n()) %>%
  ggplot(mapping = aes(x = year,
                       y = count)) +
  geom_bar(stat = 'identity') +
  scale_x_continuous(breaks = seq(1970, 2016, 2)) +
  theme(axis.text.x = element_text(angle = 30))
```

### Month

```{r}
data_c %>%
  mutate(month = month(Date, 
                       label = T)) %>%
  group_by(month) %>%
  summarise(count = n()) %>%
  ggplot(mapping = aes(x = month,
                       y = count)) +
  geom_bar(stat = 'identity') 
```

### Weekday

```{r}
data_c %>%
  mutate(wday = wday(Date, 
                     label = T)) %>%
  group_by(wday) %>%
  summarise(count = n()) %>%
  ggplot(mapping = aes(x = wday,
                       y = count)) +
  geom_bar(stat = 'identity') 
```

Obervation:
1. After 1996, the number of game released had increased very fast, and reached the highest in the year of 2008
2. Game makers prefer release their games at the end of the year and during the workdays

## Combine year and month

```{r}
data_c %>%
  filter(release_year != 1970) %>%
  mutate(month = month(Date, 
                       label = T)) %>%
  group_by(release_year,
           month) %>%
  summarise(count = n()) %>%
  ggplot(mapping = aes(x = month,
                       y = release_year,
                       fill = count)) +
  geom_tile() +
  scale_fill_gradient(low = 'white',
                      high = 'red') +
  scale_y_continuous(breaks = seq(1995, 2016, 1))
```

## Score?

```{r}
data_c %>%
  filter(release_year != 1970) %>%
  mutate(month = month(Date, 
                       label = T)) %>%
  group_by(release_year,
           month) %>%
  summarise(avg_score = mean(score)) %>%
  ggplot(mapping = aes(x = month,
                       y = release_year,
                       fill = avg_score)) +
  geom_tile() +
  scale_fill_gradient(low = 'white',
                      high = 'red') +
  scale_y_continuous(breaks = seq(1995, 2016, 1))
```

## Platfroms and Genres

```{r}
Top_5_platform <-  data_c %>%
  mutate(platform = as.character(platform)) %>%
  group_by(platform) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) %>%
  head(5) %>%
  select(platform)
Top_20_genres <- data_c %>%
  mutate(genre = as.character(genre)) %>%
  group_by(genre) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) %>%
  head(20) %>%
  select(genre)


data_c %>%
  filter((platform %in% Top_5_platform$platform) & (genre %in% Top_20_genres$genre)) %>%
  group_by(platform,
           genre) %>%
  summarise(avg_score = mean(score)) %>%
  ggplot(mapping = aes(x = platform,
                       y = genre,
                       fill = avg_score)) +
  geom_tile() +
  scale_fill_gradient(low = 'green',
                      high = 'red') +
  geom_text(aes(label = sprintf('%.2f', avg_score)))
```


